---
title: "Gov 50 Final Project"
author: "Marc P."
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project Proposal

Proposal: Analyzing Monthly Temporal Patterns of Terrorism in Western Europe (1965-2005)

Data Source:
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/HJHGYA

Research Question: How is the frequency of terrorist killings in Western European countries related to different months of the year during the period 1965-2005?

Hypothesis: Drawing from the seasonality effect observed in various socio-political phenomena, it is hypothesized that certain months might experience higher frequencies of terrorist killings compared to others. This could be due to various reasons such as favorable climatic conditions for operations, significant political or social events in specific months, or even strategic timing choices made by terrorist organizations. For instance, months with significant public events or holidays might be targeted more frequently due to the visibility and impact of such attacks.


Explanatory Variable(s) of Interest:

Month of the Year: This will be a categorical variable indicating each of the twelve months. Each terrorist killing will be tagged to a specific month based on the date of occurrence.

Outcome Variable of Interest:

Frequency of Terrorist Killings: This variable represents the number of terrorist killings in a given month in Western European countries. It will be measured as count data based on the provided dataset's information.


Expected Patterns:

Supportive Pattern: An observed consistent increase in the frequency of terrorist killings in specific months across the years would support our hypothesis. For instance, if July repeatedly has higher incidences due to national holidays or events, it would validate our assumption about strategic timing.

Contradictory Pattern: If the data shows no discernible pattern or if terrorist activities are evenly distributed across months, it would challenge our hypothesis and suggest that month-based seasonality might not play a crucial role in the timing of terrorist activities during the given period.


```{r setup, include = FALSE}
options(width = 100)
knitr::opts_chunk$set(error = TRUE)
library(dplyr)
library(knitr)
library(readr)
library(tidyverse)
library(broom)
```

```{r Analysis 1}

library(tidyverse)

load("data/DTV2009.RData")

library(dplyr)

x_clean <- x %>%
  filter(month >= 1 & month <= 12)

print(table(x_clean$month))

monthly_killings_clean <- x_clean %>%
  group_by(month) %>%
  summarize(Killings = n())

ggplot(monthly_killings_clean, aes(x = factor(month, levels = 1:12), y = Killings)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  scale_x_discrete(breaks = 1:12, labels = month.name) +
  labs(x = "Month", y = "Frequency of Terrorist Events",
       title = "Monthly Frequency of Terrorist Events in Western Europe (1965-2005)")


```

Analysis:

The bar chart visualizes the frequency of terrorist killings in Western European countries from 1965 to 2005 by month. From the chart, it's evident that May and August had the highest number of reported killings, with May being the peak. There's a noticeable increase in killings starting from April, with a peak in May, followed by a somewhat high frequency that persists through the summer and fall months before tapering off towards the end of the year. This pattern could suggest a seasonal trend in the frequency of terrorist activities, with late spring to early fall being periods of higher risk. However, to establish a solid conclusion, it would be essential to analyze the context of these events, the types of groups active during these times, and other socio-political factors that might influence these patterns

However, upon further analysis, it's important to note that this data is significantly skewed by four "mass casualty" events. Specifically, in May, there was one such event accounting for 26 deaths, which heavily influences the peak observed. In August, two mass casualty events, with 85 and 29 deaths respectively, disproportionately inflate the figures for that month. The presence of these outliers suggests that the apparent seasonal trend, with a spike in late spring and persisting through summer and fall, might not accurately reflect the general pattern of terrorist activities. For that reason, we should look at ways to potentially remove outlier events that significantly skew data.


```{r Analysis 3}

# Revised Code for Distinct Count of Attacks per Month

library(tidyverse)

load("data/DTV2009.RData")

x_clean <- x %>%
  filter(month >= 1 & month <= 12)

# Counting distinct number of attacks per month
monthly_attacks_distinct <- x_clean %>%
  group_by(month) %>%
  summarize(NumberOfAttacks = n_distinct(attack))

# Plotting the distinct number of attacks
ggplot(monthly_attacks_distinct, aes(x = factor(month, levels = 1:12), y = NumberOfAttacks)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  scale_x_discrete(breaks = 1:12, labels = month.name) +
  labs(x = "Month", y = "Distinct Number of Attacks",
       title = "Monthly Distinct Number of Terrorist Attacks in Western Europe (1965-2005)")


```

Therefore, we are now adjusting our approach. This new bar chart represents the monthly distribution of the distinct number of terrorist attacks in Western Europe from 1965 to 2005. This perspective focuses on the frequency of events rather than the severity in terms of casualties - providing us a fresh perspective that potentially removes the "mass casualty" events data skew we were faced with previously.

The new graph shows a trend where the number of distinct terrorist attacks increases through the spring, reaching an evident peak during the summer months. The small dip in the fall, particularly in August and September, followed by a rise again in October and November, suggests potential seasonal factors or operational cycles may influence terrorist activity, as our hypothesis stated.

This pattern could be related to various factors:

Hypothesis 1: Operational Conditions: Better weather in spring and summer could facilitate the logistics of carrying out attacks. Additionally, longer daylight hours may offer more opportunities for actions that require daylight.

Hypothesis 2: Vacation Periods: Summer months are typically vacation periods, which may not only affect the number of potential targets due to tourism but might also impact the availability of members within terrorist groups to both plan and execute attacks.

Hypothesis 3: Political and Social Calendar: National holidays, elections or other political events often fall within these peak periods and may serve as catalysts for terrorist activities aimed at disrupting or making a statement during these times.

The decrease in August and September could be due to a variety of factors, such as:

Hypothesis 1: Return to Routine: As people return to work and school, there may be a temporary dip in the availability of targets or in the operational capacity of terrorist networks.

Hypothesis 2: Counterterrorism Efforts: The return to work of police and national security forces from summer holidays could deter planned attacks.

Hypothesis 3: Resource Reallocation: Terrorist groups may reallocate resources after a high-activity period, leading to a temporary dip in attacks in these months.

It's important to note that while this trend is observable in the data, it's definitely only a starting point for understanding the complexities behind terrorist activities.


```{r Analysis 2}

library(ggplot2)
library(dplyr)
library(lubridate)
library(ggplot2)
library(dplyr)
library(lubridate)

x$date <- make_date(x$year, x$month, 1)

monthly_attacks <- x %>%
  group_by(date) %>%
  summarize(NumberOfAttacks = n())

ggplot(monthly_attacks, aes(x = date, y = NumberOfAttacks)) +
  geom_line(color = "steelblue") +
  geom_point(size = 0.5, color = "steelblue") +  
  geom_smooth(method = "lm", color = "darkorange") +  
  labs(x = "Time", y = "Number of Terrorist Attacks",
       title = "Monthly Number of Terrorist Attacks in Western Europe (1965-2005)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Here, we notice that the monthly number of terrorist attacks in Western Europe is steadily decreasing from 1965 to 2005. Admittedly, this graph does not provide much information in the context of our hypothesis on the seasonality of terrorist events, even though it does show that there is a notable decrease in terrorist attacks through the years - which is why I added it; things are seemingly getting better!
